LME-MPM applied to quasi-brittle fracture.

### 2020-06-22 2020_EFM_MPM_Eigensoftening.tex
Cambios MMP Conclusiones
Cambios MMP

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 881 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of
discretization were required to assess the objectiveness of the
obtained results. The results here exposed are from a discretization
of 30616 material points and 6297 nodes. To achieve better results, an
unestructured grid layout has been adopted focusing the minimum nodal
spacing, 0.47 mm, in the middle of the beam and in the impact surface
of the hammer. For this simulation, \gls{gamma-lme} was fixed to
6. Once the accuracy of the \acrshort{lme} approximants have been
proven in Section \ref{sec:3.1}, this discretization has been adopted.
%without comparing
%with linear interpolation. Therefore \\

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam.
Despite the observed discrepancies, this is out of the scope of this
document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more
realistic when modelled by a suitable contact algorithm instead of
avoiding any vertical displacement}.\MMP{Navas {\it et al.}
(2017)\cite{Navas_2017_ES} demonstrate the importance of the stiffness
of the support in order to obtain accurate results of the reaction
forces. Interesting contact algorithms within the \acrshort{mpm}
methodology are available in the literature
\cite{Bardenhagen_Contact_2001,XZhang_Contact_2011}.}

Finally, longitudinal stress distribution and crack propagation are
depicted in Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. On the one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. On the other
hand, there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Finally, further improvements of the present research are exposed. The
algorithm would have the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced
concretes\cite{Navas_2018_ES,Ruiz2019}. In addition, the eigensoftening
approach can be extrapolated to rock fracture in the same manner the
eigenerosion was~\cite{Wang2017}, taking into account that the
volumetric stiffness remains in order to support bulk loads. Also the
benefits of \acrshort{lme} can be exploited by adapting it through the
deformation gradient \cite{Kochmann2019}. \textit{A priori} this would
enhance the localisation properties of the proposed algorithms in more
sophisticated domains.

>>>>

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 2640 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of
discretization were required to assess the objectiveness of the
obtained results. The results here exposed are from a discretization
of 21576 material points and 3943 nodes. To achieve better results, an
unestructured grid layout has been adopted focusing the minimum nodal
spacing, 1.2 mm, in the middle of the beam and in the impact surface
of the hammer. For this simulation, \gls{gamma-lme} was fixed to
1.8, this low value of the regularization parameter smooth high stress
overshots after the impact. Here the stabilization properties of the \acrshort{lme}
described in \cite{Molinos2020} becomes crucial to achieve
non-spurious results. In fact, standard linear element were tested
with unsatisfactory results since grid-crossing occurs in the hammer
head producing non realistic forces.\\

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of the impact forces was correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. A feasible explanation to this unexpected result can due to a
over-rigidity in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam.
Despite the observed discrepancies, this is out of the scope of this
document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more
realistic when modelled by a suitable contact algorithm instead of
avoiding any vertical displacement}.\MMP{Navas {\it et al.}
(2017)\cite{Navas_2017_ES} demonstrate the importance of the stiffness
of the support in order to obtain accurate results of the reaction
forces. Interesting contact algorithms within the \acrshort{mpm}
methodology are available in the literature
\cite{Bardenhagen_Contact_2001,XZhang_Contact_2011}.}

Finally, longitudinal stress distribution and crack propagation are
depicted in Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. On the one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. On the other
hand, there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0005}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0015}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0030}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Finally, further improvements of the present research are exposed. The
algorithm would have the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced
concretes\cite{Navas_2018_ES,Ruiz2019}. In addition, the eigensoftening
approach can be extrapolated to rock fracture in the same manner the
eigenerosion was~\cite{Wang2017}, taking into account that the
volumetric stiffness remains in order to support bulk loads. Also the
benefits of \acrshort{lme} can be exploited by adapting it through the
deformation gradient \cite{Kochmann2019}. \textit{A priori} this would
enhance the localisation properties of the proposed algorithms in more
sophisticated domains. And the implementation of a contact algorithm to model other impact cases such as bullets collisions.

### 2020-06-22 _region_.tex
Cambios MMP

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 881 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of
discretization were required to assess the objectiveness of the
obtained results. The results here exposed are from a discretization
of 30616 material points and 6297 nodes. To achieve better results, an
unestructured grid layout has been adopted focusing the minimum nodal
spacing, 0.47 mm, in the middle of the beam and in the impact surface
of the hammer. For this simulation, \gls{gamma-lme} was fixed to
6. Once the accuracy of the \acrshort{lme} approximants have been
proven in Section \ref{sec:3.1}, this discretization has been adopted.
%without comparing
%with linear interpolation. Therefore \\

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of both forces are correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. On the other hand, if the delay is computed analytically through
concrete wave speed is much closer to the numerical results. A
feasible explanation to this unexpected result can due the composite
nature of the concrete. Another discrepancy is in the much higher
values of maximum impact and reaction forces in numerical than in
experimental. In this case, this is a consequence of over-rigidity
in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam. %In the
%%\acrshort{mpm}, the implementation of this feature can be bypassed to obtain preliminary results. Nevertheless, this kind of algorithm could
%be implemented to get ever more accurate solutions. 
Despite the observed discrepancies, this is
out of the scope of this document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more realistic when modelled by a suitable contact algorithm instead of avoiding any vertical displacement}.\\

Finally, longitudinal stress distribution and crack propagation are
depicted in Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. On the one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. On the other
hand, there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0060}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0080}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0100}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\DM{Cometario general: Pienso que hay que estructurar un poco la conclusión. No agregar tantas citas. Acá estás concluyendo tu trabajo. Iría de lo general a lo particular. Pondría claramente en cada parrafo cual es el aporte. Con tantas citas no queda claro que hiciste vos y que los otros. Las citas estarán puestas cuando vos comenzaste a hablar de lo que hiciete en el cuerpo del documento. Acá las citas minimas y tus aportes.
Imagino un primer parrafo asi diciendo cuales son las limitaciones ...propuesta modificable...:}

\DM{The paper proposes the eigensoftening algorithm in a \acrshort{MPM} framework for quasi-brittle materials to overcome the limitations observed in the EigenMPM \cite{Zhang_EE_2020}. The overestimation of tensile 
stress and strain peaks in the fracture mechanism of quasi-brittle materials and the stress oscillations observed with conventional \acrshort{mpm}/\acrshort{gimp} are faced.  }

\PNA{Based on the EigenMPM approach to fracture proposed by Zhang {\it et al.} 
\cite{Zhang_EE_2020}, two additional improvements have been proposed in the present research in order to 
solve the shortcomings of the original concept that Zhang {\it et al.} pointed out in the aforementioned paper \cite{Zhang_EE_2020}. }

Firstly, by
introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrshort{fem}. Furthermore, this failure criterion admits the design of more elaborated damage curves, which can be employed in order to describe the failure process in more sophisticated
materials. In this direction, an interesting future improvement of the
algorithm would be the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced concretes\cite{Ruiz2019}. Interesting researches have been done on this field~\cite{Navas_2018_ES}. In addition, the eigensoftening approach can be extrapolated to rock fracture in the same manner the eigenerosion was~\cite{Wang2017}, taking into account that the volumetric stiffness remains in order to support bulk loads.

Secondly, \acrshort{lme} is employed to remove from
the solution those wiggles provoked by inaccuracies in the stress integration. This spatial discretization was previously employed within the \acrshort{mpm} framework with excellent results under the elastic regime~\cite{Wobbes2020,Molinos2020}. The traditional time scheme of the \acrshort{mpm} has been improved with a Newmark Predictor Corrector scheme as well. In this research, the aforementioned techniques have helped to improve the
results in fracture propagation through the eigensoftening and eigenerosion algorithms. Further improvement may be reached by adapting \acrshort{lme} through the
deformation gradient \cite{Kochmann2019}. \textit{A priori} this would enhance the localisation properties of the proposed algorithms in more sophisticated domains.

Finally, for better capturing the measurement of forces made in the
laboratory, more realistic contact tools have to be
developed. Navas~\textif{et al.} demonstrate the importance of the
stiffness of the support in order to obtain accurate results of the
reaction forces. Interesting contact algorithms within the
\acrshort{mpm} methodology are available in the literature
\cite{Bardenhagen_Contact_2001,XZhang_Contact_2011}

@@

\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1295) }

>>>>

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2009,Zhang_2010a}. As in the aforementioned study,
an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 2640 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of
discretization were required to assess the objectiveness of the
obtained results. The results here exposed are from a discretization
of 21576 material points and 3943 nodes. To achieve better results, an
unestructured grid layout has been adopted focusing the minimum nodal
spacing, 1.2 mm, in the middle of the beam and in the impact surface
of the hammer. For this simulation, \gls{gamma-lme} was fixed to
1.8, this low value of the regularization parameter smooth high stress
overshots after the impact. Here the stabilization properties of the \acrshort{lme}
described in \cite{Molinos2020} becomes crucial to achieve
non-spurious results. In fact, standard linear element were tested
with unsatisfactory results since grid-crossing occurs in the hammer
head producing non realistic forces.\\

First, the reaction and impact forces are validated against their
experimental counterparts. Since the impact forces applied by the
hammer and the reaction forces at the two supports were experimentally
measured, they are compared with numerical ones in Figure
\ref{fig:Reactions-Forces-impact-test}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figure-impact-test-Forces-Time}
  \caption{Evolution of the reaction forces plotted \textit{versus}
    the imposed displacement for linear interpolation technique and
    the \acrshort{lme} approximants.}
  \label{fig:Reactions-Forces-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Note that the general trend of the impact forces was correctly
captured. Observe that in experimental solution the delay between
impact peak and reactions peak does not agree with the numerical
ones. A feasible explanation to this unexpected result can due to a
over-rigidity in the supports. \PNA{Although the traditional \acrshort{mpm} implementation allows collision of two bodies, inaccurate energy conservation is obtained without the implementation of a conservative contact algorithm to reproduce the hammer impact in the beam.
Despite the observed discrepancies, this is out of the scope of this
document, although future research of authors will overcome this
limitation. Similarly, results of support forces will be more
realistic when modelled by a suitable contact algorithm instead of
avoiding any vertical displacement}.\MMP{Navas {\it et al.}
(2017)\cite{Navas_2017_ES} demonstrate the importance of the stiffness
of the support in order to obtain accurate results of the reaction
forces. Interesting contact algorithms within the \acrshort{mpm}
methodology are available in the literature
\cite{Bardenhagen_Contact_2001,XZhang_Contact_2011}.}

Finally, longitudinal stress distribution and crack propagation are
depicted in Figure \ref{fig:Stress-vs-damage-impact-test}. Stress
evolution agrees with the results obtained by \cite{Navas_2017_ES}
under the \acrshort{otm} framework. Two additional considerations about
this results are concerning the interpolation technique. On the one hand,
in the opposite to the \acrshort{ugimp} shape function, \acrshort{lme}
allows unstructured grids. This is a remarkable benefit for this kind
of simulations since the majority of the computational effort is
clearly localised in a tiny region of the domain. On the other
hand, there is the possibility to reduce the shape function support to avoid
a material point in one side of the crack is affected by those in the
other side of the crack. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\centering
\subfigure[t = 0.27 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0005}}
\subfigure[t = 0.36 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0015}}
\subfigure[t = 0.45 ms]{\includegraphics[width=0.83\textwidth]{./Figure-impact-test-Damage-vs-Stress-0030}}
\caption{Evolution of the stress tensor magnitude for linear
  interpolation (pictures in the left side), and \acrshort{lme}
  (pictures in the right side). Both simulations performed with an
  eigenerosion algorithm.}
\label{fig:Stress-vs-damage-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\DM{The paper proposes the eigensoftening algorithm in a \acrshort{MPM} framework for quasi-brittle materials to overcome the limitations observed in the EigenMPM \cite{Zhang_EE_2020}. The overestimation of tensile 
stress and strain peaks in the fracture mechanism of quasi-brittle
materials and the stress oscillations observed with conventional
\acrshort{mpm}/\acrshort{gimp} are faced.} \PNA{Based on the
EigenMPM approach to fracture, two additional improvements have been proposed in the present research in order to solve the shortcomings of the original concept pointed out in the aforementioned paper \cite{Zhang_EE_2020}.}

Firstly, by introducing the concept of eigensoftening to the \acrshort{mpm} framework,
which allows to consider gradual failure process of each material point. This concept is analogous the softening law which is traditionally employed in the context
of cohesive fracture in \acrshort{fem}. Furthermore, this failure
criterion admits the design of more elaborated damage curves, which
can be employed in order to describe the failure process in more
sophisticated materials. Secondly to achieve better results, three new
numerical algorithms has been implemented. The \acrshort{lme}
approximants are employed to remove from
the solution those wiggles provoked by inaccuracies in the stress
integration. The traditional time scheme of the
\acrshort{mpm} has been improved with a Newmark Predictor Corrector
scheme as well. And the traditional concept of element connectivity
has been removed to allow a more flexible and efficient implementation of search
algorithm. In this research, the aforementioned techniques have helped
to improve the results in fracture propagation through the
eigensoftening and eigenerosion algorithms.

Finally, further improvements of the present research are exposed. The
algorithm would have the ability to simulate of crack patterns in
composite material such as steel or carbon fiber reinforced
concretes\cite{Navas_2018_ES,Ruiz2019}. In addition, the eigensoftening
approach can be extrapolated to rock fracture in the same manner the
eigenerosion was~\cite{Wang2017}, taking into account that the
volumetric stiffness remains in order to support bulk loads. Also the
benefits of \acrshort{lme} can be exploited by adapting it through the
deformation gradient \cite{Kochmann2019}. \textit{A priori} this would
enhance the localisation properties of the proposed algorithms in more
sophisticated domains. And the implementation of a contact algorithm to model other impact cases such as bullets collisions.

@@

\message{ !name(2020_EFM_MPM_Eigensoftening.tex) !offset(-1300) }
